Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors That Have Mixed During Cooling

### 2019-06-17 paper/ms.tex
Finish cleanup in results section

\section{Methodology}

@@

We frame the presentation of the results of the suites of simulations
principally in terms of the yield of \Ni{56}, the energy source of
the light curve of an event. \Ni{56} thus serves as a proxy
for the brightness of an explosion and comparison of the yields
is equivalent to comparing the brightness of events. The
yields were estimated from $Y_e$ and the NSE progress
variable, by assuming the composition in NSE is \Ni{56} plus equal
parts \Fe{54} and \Ni{58}, as described in
\citet{townetal2009,Meaketal09}.
This assumption allows the fraction of IGEs in the form of $^{56}$Ni to be estimated from the $Y_e$ tracked by the burning model.

@@

The expected final production of \Ni{56} as a function of simulation time
for 10 simulations of explosions from each progenitor is shown in Figure \ref{fig:nithybrid}.
On average, the CO model reaches the DDT phase sooner then the
hybrid model.
A few simulations ended early, but after reaching their final yields, due to problematic physical conditions in the fluff.
{\color{red}TODO: confirm.}

@@

Figure \ref{fig:masshighdens} compares the final yield of IGEs
for the C-O and C-O-Ne simulation suites with the amount of expansion of the
WD during the deflagration phase. The amount of expansion is
characterized by the mass above $2 \times 10^7~\grampercc$ at the
time of the first DDT occurrence, with more high-density mass
indicating less expansion during deflagration. The averages of both
the C-O and C-O-Ne suites along both axes are indicated by the shaded
regions with $\pm1\sigma$ widths.
Lighter shaded regions indicate the averages and standard deviations found by \citet{willcoxetal2016} for hybrid C-O-Ne progenitors that do not mix during cooling.

As expected from previous studies \citep[e.g.][]{townetal2009}, the trend for both C-O and C-O-Ne
models is that less expansion during the deflagration phase results in
greater IGE yields.
This is because less expansion allows the detonation to consume more high density fuel, which is burned more completely \citep{SeitenzahlTownsley2017}.
At moderate expansion, where the mass at high density is between around 1.0 to 1.1 M$_\odot$, the C-O models tend to yield greater IGE mass.
Following
\citet{willcoxetal2016}, we interpret
the lower IGE yield in C-O-Ne models as resulting
from the lower \C{12} abundance and the fact that, given
similar fuel density, the \Ne{20}-rich fuel will burn to cooler
temperatures than fuel in the C-O models. The result is
slower burning to IGE and thus a lower IGE yield.
At lower degrees of expansion, where there is around 1.2~M$_\odot$ of material at high densities, there appears to be little difference between the IGE yields of C-O and C-O-Ne progenetiors.
This convergence for dense, weakly expanded, cases was not noted in \citet{willcoxetal2016} because they had very few cases with more than 1.1~M$_\odot$.
It does appear consistent with an extrapolation of their data and the two cases that did yield these higher masses at high density.

This convergence of IGE yield also appears strongly in comparing the averages over the whole set found in this study and that of \citet{willcoxetal2016}.
As seen from the shaded regions in Figure \ref{fig:masshighdens} (lighter shade is unmixed), the degree of expansion seen for the our mixed progenitors is significantly less than for the unmixed cases.
This results in more mass at high density in the mixed cases, leading to more similar IGE yields.
The most likely reason for the weaker expansion is the higher central density necessary to ignite carbon burning with the lower central carbon fraction resulting from mixing during cooling \citep{brooksetal2017}.
The increase in the IGE yield that results from this higher central density is negligible for the C-O case.
This is consistent with the results of \citet{Kruegeretal2012}, who found that IGE yield is fairly independent of central density.
The C-O-Ne case, however, appears to act very differently.
The steepness of the relation between IGE yield and mass at high density is such that the mean IGE yield found here is more than 0.15~M$_\odot$ larger than that found in \citet{willcoxetal2016}.
This increase closes the gap, making the resulting IGE yields similar.

While the higher central density makes the IGE yields quite similar, the hybrid C-O-Ne WD still produces less $^{56}$Ni than a C-O WD at similar central densitiy.
In order to understand this we looked more carefully at how the process of electron capture during the explosion proceeds in our simulations.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/ni56_v_ige.pdf}
\caption{\label{fig:conversion}
Estimated yield of \Ni{56} as a function of final mass burned to IGE
for C-O-Ne progenitors (blue) and C-O progenitors (red) at
$t = 4.0 \second$.  Lines indicate corresponding relations found in W16 with
a lower central density and CONe progenitor whose interior is not mixed
during cooling.
}
\end{figure}
Figure \ref{fig:conversion} shows the estimated \Ni{56} yields
across the range of masses burned to IGE
for all C-O and C-O-Ne realizations.
The difference between the two sets of results is more pronounced
than the difference between hybrid and traditional C/O models
found by \citet{willcoxetal2016}. This difference follows from
differences in the electron capture rates, which follow from
{\color{red} two} reasons. {\color{red} HIGHER CENTRAL DENSITY?
MORE METALLICIY? HELP!}
%The conversion of \Ni{56} to IGEs are more pronounced in my case. There are two assumptions as to why figure 6 is more pronounced in my paper then in Dons.
%First is that higher critical density leads to higher electron capture resulting in less N\Ni{56}i56 produced. Figure 4 shows that CO model has a stronger electron capture then the hybrid model. The second is that the central ignition leads to early burning at higher density. Both of these cases lead to more electron capture resulting in a larger difference in evolution between the models.
%NEW: Production of 56Ni and mass burned to IGEs for CO (red) and
%hybrid CONe (green) WD realizations. SAME AS DON'S FIG 18

\begin{figure}
\includegraphics[width=\columnwidth]{figures/RatioNi56IGE_v_time_plot.pdf}
\caption{\label{fig:compare_ratio}
Plot of estimated \Ni{56} mass vs.\ time showing the evolution
for both suites of simulations. Results from
C-O-Ne progenitors are in red and C-O progenitors blue.
%SAME AS DON'S FIGURE 19
}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{figures/MBTI_v_time_plot.pdf}
\caption{\label{fig:compare_burned}
Plot of estimated mass burned to IGEs vs.\ time showing the evolution
for both suites of simulations. Results from C-O-Ne progenitors are
in red and C-O progenitors blue.
Note that one of the C/O simulations presented did not complete.
%SAME AS DON'S FIGURE 20
}
\end{figure}

Figure \ref{fig:compare_ratio} shows the fraction by mass of IGE
material producing \Ni{56} as it evolves in time and Figure
\ref{fig:compare_burned} shows the concurrent evolution of mass
burned to IGE for both suites. While there is considerable
scatter in the results, differences are obvious.

During the deflagration phase, the C-O models
on average burn more material to IGE and also had a significantly higher
fraction of IGE material that is \Ni{56}, yielding more \Ni{56} than
the C-O-Ne models. This result is counter to that observed in
\citet{willcoxetal2016}. During the detonation phase, the differences
are not as obvious and the final mass of IGEs are comparable. The
fraction of IGE material that is \Ni{56} at the end of the evolution,
however, is significantly higher for C/O models than for our hybrid models.
{\color{blue}This result differs from the results of
  \citet{willcoxetal2016} as well and is consistent with the
  observation that our hybrid models on average expand less during the
  deflagration stage than our C/O models. By contrast, the hybrid models of
  \citet{willcoxetal2016} expanded very slightly more
  than C/O models on average during the deflagration stage.}

We attribute these differences as following from the different
compositions of the progenitors. The C-O-Ne hybrid progenitors
had more neutron-rich material (a lower $Y_e$) than the
traditional C-O models. As expected, when this material burns
to IGEs, the proportion of neutron-rich material only gets
higher, as the results demonstrate.

The difference between these results and those of
\citet{willcoxetal2016} follows from their hybrid C-O-Ne progenitors
having a different temperature structure that
led to ignition of the explosion at lower densities.
The C-O models in the \citet{willcoxetal2016} study
thus had more material burning at high densities and thus more
neutronization and a lower relative \Ni{56} yield than the
corresponding hybrid C-O-Ne models. In the case of the comparison here, the
C-O and hybrid C-O-Ne models {\color{red} have the same central
temperature and density and very similar temperature and density
profiles at high densities, so the simulations in both cases
began with central ignitions. Differences in the explosion outcomes
for our C/O and hybrid C-O-Ne WD models
accordingly follow from differences in their composition.}

>>>>

\section{Methodology}
\label{sec:method}

@@

We frame the presentation of the results of the suites of simulations
principally in terms of the yield of \Ni{56}, the energy source of
the light curve of an event. \Ni{56} thus serves as a proxy
for the brightness of an explosion and comparison of the yields
is equivalent to comparing the brightness of events. The
yields were estimated from $Y_e$ and the NSE progress
variable, by assuming the composition upon NSE freeze-out is \Ni{56} plus equal
parts \Fe{54} and \Ni{58} \citep{townetal2009,Meaketal09}.
This assumption allows the fraction of IGEs in the form of $^{56}$Ni to be estimated from the $Y_e$ tracked by the burning model.

@@

The expected final production of \Ni{56}, after NSE freeze-out, as a function of simulation time
for 10 simulations of explosions from each progenitor is shown in Figure \ref{fig:nithybrid}.
On average, the CO model reaches the DDT phase sooner then the
hybrid model.
A few simulations ended early, but after reaching their final yields, due to problematic physical conditions in the fluff.
{\color{red}TODO: confirm.}

@@

Figure \ref{fig:masshighdens} compares the final yield of IGEs
for the C-O and C-O-Ne simulation suites with the amount of expansion of the
WD during the deflagration phase. The amount of expansion is
characterized by the mass above $2 \times 10^7~\grampercc$ at the
time of the first DDT occurrence, with more high-density mass
indicating less expansion during deflagration. The averages and standard deviations, $\sigma$, of both
the C-O and C-O-Ne suites along both axes are indicated by the shaded
regions with $\pm1\sigma$ widths.
Lighter shaded regions indicate the averages and standard deviations found by \citet{willcoxetal2016} for hybrid C-O-Ne progenitors that do not mix during cooling.

As expected from previous studies \citep[e.g.][]{townetal2009}, the trend for both C-O and C-O-Ne
models is that less expansion during the deflagration phase results in
greater IGE yields.
This is because less expansion allows the detonation to consume more high density fuel, which is burned more completely \citep[See discussion in][and references therein]{SeitenzahlTownsley2017}.
At moderate expansion, where the mass at high density is between around 1.0 to 1.1 M$_\odot$, the C-O models tend to yield greater IGE mass.
Following
\citet{willcoxetal2016}, we interpret
the lower IGE yield in C-O-Ne models as resulting
from the lower \C{12} abundance and the fact that, given
similar fuel density, the \Ne{20}-rich fuel will burn to cooler
temperatures than fuel in the C-O models. The result is
slower burning to IGE and thus a lower IGE yield.
At lower degrees of expansion, where there is around 1.2~M$_\odot$ of material at high densities, there appears to be little difference between the IGE yields of C-O and C-O-Ne progenetiors.
This convergence for dense, weakly expanded, cases was not noted in \citet{willcoxetal2016} because they had very few cases with more than 1.1~M$_\odot$.
It does appear consistent with an extrapolation of their data and the two cases that did yield these higher masses at high density.

This convergence of IGE yield also appears strongly in comparing the averages over the whole set found in this study and that of \citet{willcoxetal2016}.
As seen from the shaded regions in Figure \ref{fig:masshighdens} (lighter shade is unmixed), the degree of expansion seen for the our mixed progenitors is significantly less than for the unmixed cases.
This results in more mass at high density in the mixed cases, leading to more similar IGE yields.
The most likely reason for the weaker expansion is the higher central density necessary to ignite carbon burning with the lower central carbon fraction resulting from mixing during cooling
(\citealt{brooksetal2017}; Section \ref{sec:method} above).
The increase in the IGE yield that results from this higher central density is negligible for the C-O case.
This is consistent with the results of \citet{Kruegeretal2012}, who found that IGE yield is fairly independent of central density.
The C-O-Ne case, however, appears to act very differently.
The steepness of the relation between IGE yield and mass at high density is such that the mean IGE yield found here is more than 0.15~M$_\odot$ larger than that found in \citet{willcoxetal2016},
shown as lines for comparison.
This increase closes the gap, making the resulting IGE yields similar.

While the higher central density makes the IGE yields quite similar, the hybrid C-O-Ne WD still produces less $^{56}$Ni than a C-O WD at similar central densitiy.
In order to understand this we looked more carefully at how the process of electron capture during the explosion proceeds in our simulations.
While the mixed and unmixed progenitors have similar carbon fractions, the mixed progenitor explodes with a higher central density, as already mentioned.
Another difference is that the mixed progenitor here does not show the off-center shell ignition that was found previously for the unmixed progenitor.
Both of these appear to contribute to the difference in the final $^{56}$Ni yield.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/ni56_v_ige.pdf}
\caption{\label{fig:conversion}
Estimated yield of \Ni{56} as a function of final mass burned to IGE
for C-O-Ne progenitors (blue) and C-O progenitors (red) at
$t = 4.0 \second$.  Lines indicate corresponding relations found in W16 with
a lower central density and CONe progenitor whose interior is not mixed
during cooling.
}
\end{figure}
Figure \ref{fig:conversion} shows the estimated \Ni{56} yields
across the range of masses burned to IGE
for all C-O and C-O-Ne realizations.
The difference between the two sets of results is more pronounced
than the difference between hybrid and traditional C/O models
found by \citet{willcoxetal2016}.
The results found here, in which the carbon-poor compositions have a lower fractional $^{56}$Ni yield, are more consistent with the expectation for a lower peak burning temperature with lower C abundance.
This movivates a closer inspection of how the $^{56}$Ni/IGE ratio evolves during the explosion.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/RatioNi56IGE_v_time_plot.pdf}
\caption{\label{fig:compare_ratio}
Expected ratio of \Ni{56} to total IGE production by mass vs.\ time showing the evolution
for both suites of simulations. Results from C-O-Ne progenitors are in blue and C-O progenitors in red.
%SAME AS DON'S FIGURE 19
}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{figures/MBTI_v_time_plot.pdf}
\caption{\label{fig:compare_burned}
Plot of estimated mass burned to IGEs vs.\ time showing the evolution
for both suites of simulations. Results from C-O-Ne progenitors are
in blue and C-O progenitors red.
Note that one of the C/O simulations presented did not complete.
%SAME AS DON'S FIGURE 20
}
\end{figure}

The fraction, by mass, of IGE material expected to be in the form of $^{56}$Ni after freeze-out is shown in Figure \ref{fig:compare_ratio} as it evolves in time.
Also shown in Figure~\ref{fig:compare_burned} is the concurrent evolution of mass burned to IGE for both suites.
The total IGE has a mostly similar evolution, but the ratio demonstrates important differences between C-O and C-O-Ne cases.

During the deflagration phase, the C-O models
on average burn slightly more material to IGE and also had a significantly higher
fraction of IGE material that is \Ni{56}, yielding more \Ni{56} than
the C-O-Ne models. This result is counter to that observed in
\citet{willcoxetal2016}.
In that work, with a lower density progenitor with a deflagration ignited off-center, the C-O mostly showed a smaller fraction of IGE in the form of $^{56}$Ni through most of the deflagation phase.
That difference was then closed during the detonation phase, to obtain similar ratios as seen in the lines above in Figure~\ref{fig:conversion}.
A similar effect happens here during the detonation phase, but the fraction of material that is $^{56}$Ni stays higher for the C-O progenitors.

The differences seen in the deflagration phase in Figure~\ref{fig:compare_ratio} appear to be critical.
Central ignition at a higher central density leads to much more electron capture during the deflagration phase for the C-O-Ne case.
Remarkably, up to around 0.5 seconds, no $^{56}$Ni is expected to be produced by the C-O-Ne progenitor.
The comparison to the results of \citep{willcoxetal2016} indicate that the central ignition, in higher density regions, is important to this ignition.
Comparing to the C-O progenitor here indicates that the lower energy release may also slow the rise of plumes, allowing more electron capture.

In conclusion, we attribute the following major differences from the progenitor treated here, which is mixed during cooling and centrally ignited, compared to the case in \citet{willcoxetal2016} that did not mix while cooling:
The higher central density necessitated for ignition by the lower central carbon fraction leads to a higher production of IGE.
The higher central density also leads to stronger electron capture.
The off-center ignition used for the C-O-Ne progenitor in \citet{willcoxetal2016} appears to have suppressed electron capture due to the deflagration being more isolated to lower densities.
Here we find that electron capture is noticably enhanced in the C-O-Ne case compared to C-O at the same density and ignition location.

These features together indicate that explosions from C-O-Ne progenitors should have similar IGE yields to normal C-O progenitors but lower kinetic energies.
This is due to their being higher central density, and therefore more tightly bound.
At the same time, C-O-Ne explosions should show lower $^{56}$Ni yield due to enhanced electron capture during the deflagration.
